---
title: "Credit Card Fraud Model Evaluation"
author: "Predicting Credit Card Fraud with R"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(dplyr)
library(ggplot2)
root <- normalizePath("..")
metrics_path <- file.path(root, "models", "model_metrics", "comparison.csv")
json_path <- file.path(root, "models", "model_metrics", "metrics.json")
sample_path <- file.path(root, "data", "processed", "scored_sample.csv")
```

## Overview

This report summarizes fraud-classification results on a highly imbalanced credit-card transaction dataset.
Accuracy is intentionally **not** used as the primary metric. Evaluation focuses on **precision, recall, F1, ROC-AUC, PR-AUC**, and false negatives.

The model supports human review and does not permanently block cards.

```{r load}
if (!file.exists(metrics_path)) {
  cat("Run `Rscript scripts/run_pipeline.R` first to generate metrics.")
  knitr::knit_exit()
}
cmp <- read.csv(metrics_path)
meta <- jsonlite::fromJSON(json_path)
```

## Selected model

- **Model:** `r meta$selected_model`
- **Version:** `r meta$model_version`
- **Decision threshold:** `r meta$threshold`
- **Balance method (train only):** `r meta$balance_method`

### Test-set metrics

```{r test-metrics}
knitr::kable(as.data.frame(meta$test), digits = 4)
```

## Model comparison (validation, optimized threshold)

```{r comparison}
knitr::kable(cmp, digits = 4)
```

```{r plot-f1, fig.width=8, fig.height=4}
val <- cmp[cmp$dataset == "validation_optimized", ]
ggplot(val, aes(x = reorder(model, f1), y = f1, fill = model)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(title = "Validation F1 at optimized threshold", x = NULL, y = "F1") +
  theme_minimal()
```

```{r plot-pr, fig.width=8, fig.height=4}
ggplot(val, aes(x = reorder(model, recall), y = recall, fill = model)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(title = "Validation recall (fraud detection rate)", x = NULL, y = "Recall") +
  theme_minimal()
```

## Score distribution

```{r scores, fig.width=8, fig.height=4}
if (file.exists(sample_path)) {
  sc <- read.csv(sample_path)
  ggplot(sc, aes(x = fraud_probability, fill = Class)) +
    geom_histogram(bins = 40, alpha = 0.7, position = "identity") +
    labs(title = "Fraud probability on held-out sample", x = "P(fraud)", y = "Count") +
    theme_minimal()
}
```

## Limitations

- Features `V1`–`V28` are anonymized PCA components; do not invent business meanings.
- Public / synthetic data is not a production bank feed.
- SMOTE/ROSE applied only to training data to avoid leakage.
